### 2020-10-05 appendixB/appendixB.Rmd
Creating markdown to show results of seed dispersal

---
title: "Simulation of seed dispersal distance"
author: "Javiera Rudolph"
output: 
  html_document:
    toc: true
    fig_width: 12
    fig_height: 6
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = FALSE, message = FALSE, fig.cap = " ")

```

```{r loadPackages}
library(tidyverse)
library(lubridate)
library(cowplot)
```

```{r importData}
seed_data_popind <- readRDS("appendixA/seed_dispersal_popind.RDS")

seed_data_popfam <- readRDS("appendixA/seed_dispersal_popfam.RDS")

```