### 2020-10-12 appendixB/sim10/appendixB_10.Rmd
arranged all tables according to ldd n

---
title: "Simulation of seed dispersal distance - 10 runs"
author: "Javiera Rudolph"
output: 
  html_document:
    toc: true
    fig_width: 12
    fig_height: 6
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(echo = FALSE, message = FALSE, fig.cap = " ")

```

```{r loadPackages}
library(tidyverse)
library(lubridate)
library(cowplot)
```

```{r importData}

directory <- "appendixB/sim10/"

seed_data_popind <- readRDS(paste0(directory, "seed_dispersal_popind.RDS"))

seed_data_popfam <- readRDS(paste0(directory, "seed_dispersal_popfam.RDS"))

```

# Seed dispersal distance at the different organizational levels

```{r}
seed_data_popind %>%
  mutate(., type = ifelse(data == "POP", "POP", "IND")) %>% 
  #filter(., dist == "exp") %>% 
  ggplot(., aes(x = data, y = dispersal, color = type)) +
  facet_wrap(~dist, ncol = 4) +
  geom_boxplot() -> a

seed_data_popfam %>%
  mutate(., type = ifelse(data == "POP", "POP", "FAM")) %>% 
  #filter(., dist == "exp") %>% 
  ggplot(., aes(x = data, y = dispersal, color = type)) +
  facet_wrap(~dist, ncol = 4) +
  geom_boxplot() -> b

plot_grid(a, b, nrow = 2)
```

```{r}

seed_data_popind %>%
  mutate(., type = ifelse(data == "POP", "POP", "IND")) %>% 
  #filter(., dist == "exp") %>% 
  ggplot(., aes(x = dispersal, color = type, group = data)) +
  facet_wrap(~dist, ncol = 4) +
  geom_density() +
  scale_x_log10() +
  NULL -> a

seed_data_popfam %>%
  mutate(., type = ifelse(data == "POP", "POP", "FAM")) %>% 
  #filter(., dist == "exp") %>% 
  ggplot(., aes(x = dispersal, color = type, group = data)) +
  facet_wrap(~dist, ncol = 4) +
  geom_density() +
  scale_x_log10() +
  NULL -> b

plot_grid(a, b, nrow = 2)

```

The figure I actually want should have all these box plots on the same one because I'm still missing the mixed distribution one.

```{r}

mixed_ind <- seed_data_popind %>% 
  mutate(type = ifelse(data == "POP", "POP", "IND")) %>%
  filter(., type == "IND") %>% 
  filter(., bic_b == "1") %>% 
  mutate(type = "BIC") %>% 
  bind_rows(., seed_data_popind %>% 
  mutate(type = ifelse(data == "POP", "POP", "IND")) %>%
  filter(., type == "IND") %>% 
  filter(., aic_b == "1") %>% 
  mutate(type = "AIC")) %>% 
  mutate(dist = "mixed")

seed_data_popind %>% 
  mutate(type = ifelse(data == "POP", "POP", "IND")) %>% 
  bind_rows(., mixed_ind) %>% 
  ggplot(., aes(x = dist, y = dispersal, color = type)) +
  #facet_wrap(~dist, scales = "free") +
  geom_boxplot() -> a

mixed_fam <- seed_data_popfam %>% 
  mutate(type = ifelse(data == "POP", "POP", "FAM")) %>%
  filter(., type == "FAM") %>% 
  filter(., bic_b == "1") %>% 
  mutate(type = "BIC") %>% 
  bind_rows(., seed_data_popfam %>% 
  mutate(type = ifelse(data == "POP", "POP", "FAM")) %>%
  filter(., type == "FAM") %>% 
  filter(., aic_b == "1") %>% 
  mutate(type = "AIC")) %>% 
  mutate(dist = "mixed")

seed_data_popfam %>% 
  mutate(type = ifelse(data == "POP", "POP", "FAM")) %>% 
  bind_rows(., mixed_fam) %>% 
  ggplot(., aes(x = dist, y = dispersal, color = type)) +
  #facet_wrap(~dist, scales = "free") +
  geom_boxplot() -> b

plot_grid(a, b, nrow = 2)

```

# Summary stats

Individual level

```{r}
seed_data_popind %>% 
  mutate(type = ifelse(data == "POP", "POP", "IND")) %>% 
  bind_rows(., mixed_ind) %>% 
  mutate(., LDD = ifelse(dispersal >= 500, 1, 0)) %>% 
  group_by(type, dist) %>% 
  summarise(n = n(),
            mean_disp = mean(dispersal),
            max_disp = max(dispersal), 
            n_ldd = sum(LDD),
            prcnt_ldd = n_ldd/n) %>% 
  arrange(., desc(prcnt_ldd)) %>% 
  knitr::kable()

```

Family level
```{r}
seed_data_popfam %>% 
  mutate(type = ifelse(data == "POP", "POP", "FAM")) %>% 
  bind_rows(., mixed_fam) %>% 
  mutate(., LDD = ifelse(dispersal >= 500, 1, 0)) %>% 
  group_by(type, dist) %>% 
  summarise(n = n(),
            mean_disp = mean(dispersal),
            max_disp = max(dispersal), 
            n_ldd = sum(LDD),
            prcnt_ldd = n_ldd/n) %>% 
  arrange(., desc(prcnt_ldd)) %>% 
  knitr::kable()
```

```{r eval = FALSE}

# Spatial data
sims.ind <- readRDS(paste0(directory, "individualSims.RDS"))

sims.ind %>% 
  map(., "exp_move")

sims.ind$B1[[1]]$exp_move

get_seed_disp_info <- function(simulations){
  models <- c("exp_move", "gam_move", "weib_move", "lnorm_move")

  df <- NULL
  for(i in 1:4){
    oneModel <- map(simulations, models[i])
    dispInfo <- map(oneModel, "seedInfo") %>% 
      unlist()
    dispersal <- map(dispInfo, "seed_disp") %>%
      unlist()

    out <- data.frame(model = models[i], 
                      #dispersal = dispersal,
                      dispInfo)
    df <- rbind.data.frame(df, out)
  }

  return(df)
}

get_seed_disp_info(sims.ind)

IDs <- unique(seed_data_popind$data)[2:13]

indSeed_data <- NULL
indSims <- vector("list", 12)
for(i in 1:12){
  a <- indSim_fx(paramList = popind, individual = IDs[i])

  b <- get_seed_disp_info(a) %>%
    mutate(data = IDs[i])

  indSeed_data <- rbind(indSeed_data, b)

  indSims[[i]] <- a
}

```