### 2021-02-01 paper/figures_tables.Rmd
Up to figure 3 with captions

---
title: "Figures and Tables"
subtitle: ' '
output:
  html_document:
    toc: yes
  word_document:
    toc: yes
bibliography: ../paper/mybibfile.bib
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)
```

```{r libraries}
library(ggplot2)
library(aracari)
library(cowplot)
library(dplyr)
library(extRemes)

```

```{r}
load("paper/fevd.RData")
```

```{r}
# Simulate the data for one run

set.seed(5)

data(ptpl)

# Panel 2: Table with gut retention times and the gamma distribution used for it
n <- 10000
grt_dens <- data.frame(seq = 1:n, dat = (rgamma(n, shape = 4, scale = 5) + 8))
grt_data <- data.frame(time = round(sample(grt_dens$dat, 5)), seq = 1:5)
tt <- max(grt_data$time)
ggplot(grt_dens, aes(x = dat)) +
  geom_density() + 
  geom_point(data = grt_data, aes(x = time, y = 0), color = "red") +
  #geom_point(aes(x = tt, y = 0), color = "red") +
  theme_bw() + labs(x = "Time in minutes", y = "Density") -> fig_GRT

# Panel 3: Exponential distribution used for animal movement lengths
rate <- 1/mean(ptpl$mpm)
md_dens <- data.frame(seq = 1:n, dat = rexp(n, rate = rate))
md_hist <- data.frame(md = round(sample(md_dens$dat, tt)))

ggplot(md_dens, aes(x = dat)) +
  geom_density() + 
  geom_point(data = md_hist, aes(x = md, y = 0)) +
  theme_bw() + labs(x = "Distance in meters", y = "Density") -> fig_MD

ggplot(md_hist, aes(x = md)) +
  geom_histogram() +
  theme_bw() + labs(x = "Distance in meters", y = "Count") -> fig_MD_hist

# Panel 4: Animal trajectory and locations of the seeds at the time the seeds are dropped
set.seed(5)
distance <- md_hist$md
angle <- runif(tt, min = 0, max = 360)
distx <- distance*cos(angle)
xloc <- c(0, cumsum(distx))
disty <- distance*sin(angle)
yloc <- c(0, cumsum(disty))
animalTraj <- data.frame(time = 0:tt, xloc = xloc, yloc = yloc)
seed_loc <- merge(animalTraj, grt_data, by = "time")
mean_seed_loc <- data.frame(x = mean(seed_loc$xloc), y = mean(seed_loc$yloc))

ggplot() +
  geom_path(data = animalTraj, aes(x = xloc, y = yloc), size = 1, color = "darkgrey") +
  geom_point(data = animalTraj, aes(x = xloc, y = yloc), color = "black", size = 1) +
  geom_point(data = seed_loc, aes(x = xloc, y = yloc), color = "red", size = 3) +
  geom_point(aes(x = 0, y = 0), color = "black", size = 5)+
  geom_vline(aes(xintercept=0), linetype = "dotted") + geom_hline(aes(yintercept = 0), linetype = "dotted") +
  labs(x = "x", y = "y") +
  coord_fixed() +
  theme_bw() -> fig_traj

ggplot() +
  geom_point(data = seed_loc, aes(x = xloc, y = yloc), color = "red", size = 3) +
  geom_point(aes(x = 0, y = 0), color = "black", size = 5)+
  geom_vline(aes(xintercept=0), linetype = "dotted") + geom_hline(aes(yintercept = 0), linetype = "dotted") +
  geom_segment(aes(x = seed_loc$xloc, xend = 0, y = seed_loc$yloc, yend = 0), linetype = "dashed", color = "red") +
  labs(x = "x", y = "y") +
  coord_fixed() +
  theme_bw() -> fig_disp

ggplot() +
  geom_point(data = seed_loc, aes(x = xloc, y = yloc), color = "red", size = 3) +
  geom_point(aes(x = 0, y = 0), color = "black", size = 5)+
  geom_vline(aes(xintercept=0), linetype = "dotted") + geom_hline(aes(yintercept = 0), linetype = "dotted") +
  #geom_segment(aes(x = seed_loc$xloc, xend = 0, y = seed_loc$yloc, yend = 0), linetype = "dashed", color = "red") +
  geom_point(data = mean_seed_loc, aes(x = x, y = y), color = "blue", size = 4) +
  geom_segment(aes(x = seed_loc$xloc, xend = mean_seed_loc$x, y = seed_loc$yloc, yend = mean_seed_loc$y),
               linetype = "dashed", color = "blue") +
  geom_segment(aes(x = 0, xend = mean_seed_loc$x, y = 0, yend = mean_seed_loc$y),
               linetype = "solid", color = "black") +
  labs(x = "x", y = "y") +
  coord_fixed() +
  theme_bw() -> fig_agg

```

#### Figure 1. 
Density distributions from which we sample **A.** gut retention time (GRT) and **B.** movement distance (MD). Red dots represent an example for one simulation run, where we sample five gut retention times. The maximum GRT for that simulation run determines the number of movement distances sampled in that simulation run, represented in the figure by the black dots.

```{r fig.width=2.5, fig.height=4}
plot_grid(fig_GRT, fig_MD, align = "v", nrow = 2, labels = "AUTO", label_size = 10)
```

#### Figure 2.
Example of one simulation run and variables measured for each run. **A.** Shows the animal trajectory followed after starting at the origin (big circle at location 0,0). The animal moves every minute following the movement distances (MD) sampled from the distribution as show in example of Figure 1. Every small black dot represents the animal's location at every minute interval. Seeds get dropped every time the simulation reaches one of the sampled gut retention times (GRT), at the animal's location in that time (red dots). **B.** Seed dispersal calculations as euclidean distances from the parent plant located at the origin (black circle at 0,0) to each of the five seeds for the simulation run (red dots). **C.** Calculation of seed dispersion measures and average seed dispersal. The average seed location for the simulation run is shown by the blue circle. The distance from each seed to the average location is used to calculate seed dispersion as a measure of seed aggregation, dashed blue lines. Average seed dispersal is calculated as the distance from the parent plant to the average seed location, shown as the black line connecting the black circle to the blue dot.

```{r fig.height=2, fig.width=6}
plot_grid(fig_traj, fig_disp, fig_agg, ncol = 3, align = "h", labels = "AUTO", label_size = 8)
```

#### Figure 3.
Variation in Weibull and Generalized Pareto distributions determined by the value of the shape parameter. **A.** In the case of the Weibull distribution, the scale parameter is defined as $b=1$, and the shape parameters are defined as follows: solid line $a=\nu=0.4$ shows a heavy tail, the dashed line $a=\nu=2$ shows a thin tail, and $a=\nu=1$ describes an exponential tail with the dotted line. This follows the assumption that a shape parameter $\nu<1$ describes a fat tail, $\nu>1$ a thin tail, and $\nu=1$ a tail following an exponential distribution. *Check here these statements about the tail. Maybe parameterization has changed it, but I don't see the heavy tail of the solid line with $\beta > 1$* **B.** In the case of the Generalized Pareto distribution, a shape parameter $=0$ describes an exponential distribution function, shape $>0$ describes a heavy tail, and $<0$ a tail with Beta distribution function, bounded at an upper value as a function of the threshold, shape and scale parameters. In the second panel, the scale parameter is set to $\sigma = 1$ the solid line $\xi = 1$ shows a heavy tail, dashed line $\xi=-0.6$ shows a Beta tail, and $\xi = 0$ describes aan exponential decay with the dotted line.

```{r fig.width=6, fig.height=2.5}
# shape
exptail <- 1
thintail <- 2
fattail <- 0.4

#scale
b <- 1
kappaexp <- b^-exptail
kappathin <- b^-thintail
kappafat <- b^-fattail

ggplot(data = data.frame(x = c(0.1, 4)), aes(x)) +
  stat_function(fun = dweibull, n = 100, args = list(scale = b, shape = thintail),
                color = "black", linetype = "dashed", size = 0.8) +
  stat_function(fun = dweibull, n = 100, args = list(scale = b, shape = exptail), 
                color = "black", linetype = "dotted", size = 0.8) +
  stat_function(fun = dweibull, n = 100, args = list(scale = b, shape = fattail),
                color = "black", linetype = "solid", size = 0.8) +
  ylab("") + xlab("") +
  lims(y = c(0, 1)) +
  theme_bw() -> weib_dens

# The shape parameter determines the behavior of the tail of the Generalized Pareto
# ?extRemes::devd
# shape = 0 gives rise to exponential disttribution function (thin tail)
# shape > 0 is a heavy tail
# shape < 0 is a tail following a Beta distribution function, a bounded upper tail at location-scale.u/shape

# shape
exptail <- 0
heavytail <- 1.2
betatail <- -0.6

# scale
gp_scale <- 1

ggplot(data = data.frame(x = c(0.1, 4)), aes(x)) +
  stat_function(fun = devd, n = 100, args = list(scale = gp_scale, shape = betatail, type = "GP"),
                color = "black", linetype = "dashed", size = 0.8) +
  stat_function(fun = devd, n = 100, args = list(scale = gp_scale, shape = exptail, type = "GP"), 
                color = "black", linetype = "dotted", size = 0.8) +
  stat_function(fun = devd, n = 100, args = list(scale = gp_scale, shape = heavytail, type = "GP"),
                color = "black", linetype = "solid", size = 0.8) +
  ylab("") + xlab("") +
  lims(y = c(0, 1)) +
  theme_bw() -> gp_dens

plot_grid(weib_dens, gp_dens, ncol = 2, labels = "AUTO", label_size = 8)

```

## Seed dispersal distances and aggregation

```{r fig.width=6}
bind_rows(null_dispersal %>% mutate(model = "Null"), 
          indiv_dispersal %>% mutate(model = "Individual"),
          fam_dispersal %>% mutate(model = "Family")) %>% 
  mutate(model = factor(model, levels = c("Null", "Individual", "Family"))) -> dispersal_data

dispersal_data %>% 
  ggplot(., aes(x = dispersal, group = model)) +
  facet_wrap(~model, ncol = 3) +
  geom_histogram(aes(y = ..density..), binwidth = 50) +
  theme_bw() +
  theme(strip.background = element_rect(fill = "white")) +
  labs(y = "Density", x = "Seed dispersal (m)") -> histograms

null_dispersion %>% 
  mutate(model = "Null") %>% 
  bind_rows(indiv_dispersion %>% 
              mutate(model = "Individual")) %>% 
  bind_rows(fam_dispersion %>% 
              mutate(model = "Family")) %>% 
  mutate(model = factor(model, levels = c("Null", "Individual", "Family"))) -> all_disp_data

all_disp_data %>% 
  ggplot(., aes(y = mean_xy_dispersal, x = model)) +
  geom_boxplot() +
  theme_bw() +
  labs(y = "Mean dispersal (m)") +
  theme(axis.title.x = element_blank()) -> a

all_disp_data %>% 
  ggplot(., aes(y = seed_dispersion, x = model)) +
  geom_boxplot() +
  theme_bw() +
  labs(y = "Seed dispersion (m)") +
  theme(axis.title.x = element_blank()) -> b

boxplots <- plot_grid(a,b, ncol = 2, align = "h", labels = c("b", "c"))

plot_grid(histograms, boxplots, nrow=2, labels = c("a", ""))

# Edit redundant labels
# Fix names for axes

```

## Long distance dispersal

The mean dispersal in the boxplots is not the same as the dispersal distances before. This mean dispersal is the distance from the origin to the mean seed location in each simulation run. The mean seed location is used to calculate dispersion, as the average distance of each seed to the mean seed location, as shown in Figure 1.

```{r}
knitr::kable(dispersal_kernel_table %>% 
               rename(Long_distance = LDD))
```

## Weibull seed dispersal kernel 
```{r}
knitr::kable(weibull_table)
```

## Generalized Pareto tail distribution for seed dispersal

```{r}
knitr::kable(evd_table_500)
```