### 2021-02-04 paper/test_thresh_dec.Rmd
Comparing declustering, and orig vs specific thresholds

---
title: "Untitled"
author: "Javiera Rudolph"
date: "2/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
knitr::opts_chunk$set(message = FALSE, warning = FALSE, echo = FALSE)
```

```{r}

load("paper/thresholds.RData")
```

```{r}
library(dplyr)
library(ggplot2)
library(cowplot)
library(extRemes)
```

# NULL

```{r}
orig_thres <- 500

#*****************************************************************************************
# NULL MODEL

# 1. Determine threshold : done in threshold.R script
null_thresh
# 2. Fit Generalized Pareto Distribution

dec_null <- decluster(null_dispersal$dispersal, threshold = null_thresh)
null_fit_dec <- fevd(dec_null, threshold = null_thresh, type = "GP")
plot(null_fit_dec)
summary(null_fit_dec)
ci(null_fit_dec, type = "parameter")

null_fit <- fevd(null_dispersal$dispersal, threshold = null_thresh, type = "GP")
plot(null_fit)
summary(null_fit)
ci(null_fit, type = "parameter")

null_fit_500 <- fevd(null_dispersal$dispersal, threshold = orig_thres, type = "GP")
plot(null_fit_500)
summary(null_fit_500)
ci(null_fit_500, type = "parameter")
```

# INDIVIDUAL

```{r}
indiv_thresh

dec_indiv <- decluster(indiv_dispersal$dispersal, threshold = indiv_thresh)
indiv_fit_dec <- fevd(dec_indiv, threshold = indiv_thresh, type = "GP")
plot(indiv_fit_dec)
summary(indiv_fit_dec)
ci(indiv_fit_dec, type = "parameter")

indiv_fit <- fevd(indiv_dispersal$dispersal, threshold = indiv_thresh, type = "GP")
plot(indiv_fit)
summary(indiv_fit)
ci(indiv_fit, type = "parameter")

indiv_fit_500 <- fevd(indiv_dispersal$dispersal, threshold = orig_thres, type = "GP")
plot(indiv_fit_500)
summary(indiv_fit_500)
ci(indiv_fit_500, type = "parameter")
```

# FAMILY

```{r}
fam_thresh

dec_fam <- decluster(fam_dispersal$dispersal, threshold = fam_thresh)
fam_fit_dec <- fevd(dec_fam, threshold = fam_thresh, type = "GP")
plot(fam_fit_dec)
summary(fam_fit_dec)
ci(fam_fit_dec, type = "parameter")

fam_fit <- fevd(fam_dispersal$dispersal, threshold = fam_thresh, type = "GP")
plot(fam_fit)
summary(fam_fit)
ci(fam_fit, type = "parameter")

fam_fit_500 <- fevd(fam_dispersal$dispersal, threshold = orig_thres, type = "GP")
plot(fam_fit_500)
summary(fam_fit_500)
ci(fam_fit_500, type = "parameter")
```