### 2021-03-19 README.Rmd
add a link

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  eval = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# aracari <img src="images/aracari.png" align="right" height="139" />

<!-- badges: start -->
<!-- badges: end -->

Access the

- [seed dispersal paper](https://javirudolph.github.io/aracari/paper/paper.html) and the [figures](https://javirudolph.github.io/aracari/paper/figures_tables.html)
- [animal movement paper](https://javirudolph.github.io/aracari/movement_paper/rudolph_ch2.html)  
- [distributions paper](https://javirudolph.github.io/aracari/distributions_paper/rudolph_distributions.html)
  - [info on the data](https://javirudolph.github.io/aracari/distributions_paper/format_raw_dat.html)

---
## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("javirudolph/aracari")
```