### 2021-03-26 introduction/introduction.Rmd
Organizing intro files

---
title: "Introduction"
output:
  html_document:
    df_print: paged
bibliography: intro_refs.bib
csl: ecology-letters.csl
---

# Introduction

## What is dispersal?

## What is the difference with movement?

## What is the role of individual variation in dispersal and movement?

## How can we explore the role and importance of this variation?

## What statistical tools can we leverage to understand and predict the role of individual variation in dispersal and how does this extend to populations and ecosystems?

## The importance of characterizing dispersal in metapopulation models