### 2022-02-14 Ch3_samplesize/Datagen_viz.Rmd
Densities for the social group.

---
title: "Mixture distribution data generation"
author: "Javiera Rudolph"
date: "1/20/2022"
output:
  html_document:
  toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
knitr::opts_chunk$set(fig.width=6, fig.height=4)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(cowplot)
```

```{r}
set.seed(20220214)

## Functions ----------------------------------------------------
# Building this function so that we get the parameters for a desired mean and standard deviation using a lognormal distribution
desired_mean_sd <- function(mean, sd){

  mu <- log(mean*2/(sqrt(mean*2+sd*2)))
  sigsq <- log(1+(sd*2/mean*2))

  return(data.frame(mu = mu, sigsq = sigsq))
}

exp.val <- function(mu, sigsq){
  exp(mu + sqrt(sigsq))
}
```

We use a mixture model to simulate step lengths for animal movement using a lognormal distribution for step length. We incorporate two levels of heterogeneity by including social group and individual variation. The assumption follows that individuals belonging to the same group will follow a similar movement but with a different variance.

We use a mixture model since this is a common approach to characterize subpopulations, and the mixture distribution is sort of a weighted average for the contributions of each subpopulation to the general population. 
$$ 
f(x) = \sum_{i=1}^{n} \pi_{i} f_i(x)
$$
Where $\pi_i$ represents the weight associated to social group $i$ and $f_i(x)$ is the distribution describing the step lengths of the social group. We incorporate individual variation by describing the distribution of social group step lengths $f_i(x)$ as a mixture as well.

$$
f_i(x) = \sum_{j=1}^s w_j g_j(x)
$$

We assume that each individual has the same weight (i.e., $w_1 = w_2 = \cdots = w_s$), and that the proportion of individuals in each social group determines the weight for the social group (i.e., $\pi_i$) relative to the overall population.

For our toy model, we assume four social groups ($i=4$), where each group has between 4-7 individuals assigned randomly, with a total of 22 individuals in our simulated population.

```{r echo=TRUE}
# MIXTURE MODEL --------------------------------------------------
## Mixture proportions ----------------------------------------
# We assume four categories of individuals with increasing movement.

ngroup <- 4
npergroup <- sample(3:7, ngroup, replace = TRUE) 
npop <- sum(npergroup)

pis <- npergroup/npop
#sum(pis)
#hist(pis)

## Mixture components --------------------------------------
### Each of these curves represents the average of the social group.
# Since movement lengths are only positive, we use a lognormal distribution to describe them

# CASE1

pars <- desired_mean_sd(mean = c(160, 300, 600, 1000), sd = c(90, 120, 160, 200))

# Special case I am considering where for each category the mean doubles, and then sd=mean*0.6
# pars <- desired_mean_sd(mean = c(150, 300, 600, 1200), sd = c(90, 180, 360, 720))

mus <- pars$mu
sigsqs <- pars$sigsq
dens_cols <- c("#264653", "#2a9d8f", "#f4a261", "#e76f51")

### PLOT the mixture components -------------------------------
lnorm_densities <- purrr::map(1:4, function(y) stat_function(fun = dlnorm,
                                                             args = list(meanlog = mus[y], sdlog = sigsqs[y]),
                              color = dens_cols[y], size=1))
ggplot() +
  lnorm_densities +
  theme_minimal() +
  labs(y = "Density") +
  lims(x = c(0, 150)) -> densities_plot
densities_plot

```

```{r}

# SAMPLING --------------------------------------------------------------

samp.size <- 50000
all.samples <- rep(0,samp.size)
for(i in 1:samp.size){

  which.cat <- sample(1:4, size=1, replace=TRUE, prob=pis)
  all.samples[i] <- rlnorm(n=1,meanlog=mus[which.cat], sdlog=sigsqs[which.cat])
}

## PLOT the samples ---------------------------------------------
# This is our f(x) from the main text
data.tail <- data.frame(values = all.samples, y = 100) %>% arrange(desc(values)) %>% filter(values >=250)
head(data.tail)

ggplot(data.frame(all.samples), aes(x = all.samples)) +
  geom_histogram(bins = 100) +
  geom_point(data = data.tail[1:50,], aes(x = values, y = y), color = "black", alpha = 0.5) +
  labs(y = "Frequency", x = "Distance") +
  theme_minimal() -> sampleshist

# VIZ -----------------------------------
plot_grid(densities_plot, sampleshist, nrow=2, labels="AUTO")
#ggsave("Ch3_samplesize/TestFig.png")

summary(all.samples)

```