### 2022-02-15 Ch3_samplesize/mess/Ch3_toymodel_sgonly.Rmd
Trying to tackle the mess. Moved things around and renamed.

---
title: "Toy Model"
author: "Javiera Rudolph"
date: "1/20/2022"
output:
  html_document:
  toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
knitr::opts_chunk$set(fig.width=6, fig.height=4)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
```

```{r message=FALSE, warning=FALSE}
set.seed(20220201)

library(dplyr)
library(tidyr)
library(ggplot2)
library(cowplot)
```

```{r Functions}
## Functions ----------------------------------------------------
# Building this function so that we get the parameters for a desired mean and standard deviation.
desired_mean_sd <- function(mean, sd){

  mu <- log(mean*2/(sqrt(mean*2+sd*2)))
  sigsq <- log(1+(sd*2/mean*2))

  return(data.frame(mu = mu, sigsq = sigsq))
}

exp.val <- function(mu, sigsq){
  exp(mu + sqrt(sigsq))
}

### Lomax functions --------------------------

lomax.pdf <- function(x,alpha,k, log.scale=FALSE){

  if(log.scale==FALSE){out <- (k/(alpha+x))*(alpha/(alpha+x))^k
  }else{
    out <- log(k) + k*log(alpha) - (k+1)*log(alpha+x)
  }

  return(out)
}

lomax.cdf <- function(x,alpha,k){

  return(1-(alpha/(alpha+x))^k)

}

ft.nllike <- function(guess=init.betas, designmat=designmat,Y=Y){

  Xbeta         <- designmat%*%guess
  alphas        <- exp(Xbeta) # because alpha = ln(X*betas)
  n             <- length(Y)
  sumlogapy     <- sum(log(alphas+Y))
  k.hat         <- n/(sumlogapy - sum(Xbeta))
  lnft.yis     <- lomax.pdf(x=Y, alpha=alphas,k=k.hat,log.scale=TRUE)
  lnL           <- sum(lnft.yis)
  nll           <- -lnL
  return(nll)
}

lomax.glm <- function(formula=~1, my.dataf, response){

  Y           <- response
  n           <- length(Y)
  designmat   <- model.matrix(formula, data=my.dataf)
  nbetas      <- ncol(designmat)
  init.betas  <- rep(1.5,nbetas)

  opt.out <- optim(par=init.betas, fn=ft.nllike, method="BFGS",
                   designmat=designmat, Y=Y)

  mles          <- opt.out$par
  nll.hat       <- opt.out$value
  BIC.mod       <- 2*nll.hat + nbetas*log(length(Y))
  Xbeta.hat     <- designmat%*%mles
  alphas.hat    <- exp(Xbeta.hat)
  sumlogapy.hat <- sum(log(alphas.hat+Y))
  k.hat         <- n/(sumlogapy.hat - sum(Xbeta.hat))

  out.list <- list(opt.out = opt.out, designmat=designmat,Y=Y, mles=mles, nll.hat=nll.hat, BIC.mod = BIC.mod,
                   alphas.hat=alphas.hat, k.hat=k.hat,data=my.dataf)

  return(out.list)

}

# Probability that it is greater.
lomax.st <- function(x=x, alpha=alpha, k=k){
  out <- alpha/(alpha+x)
  return(out^k)
}

# calculate the mean of the fit

lomax.mean <- function(alpha=alpha, k=k){
  return(alpha/(k+1))
}

# Assess the fit using model quantiles to make a qqplot
lomax.quantile <- function(alpha, k, p){
  return(alpha*(((1-p)^(1/k))-1))
}

```

```{r}

#########  playing with a simulation trial to assess how good hat(P(X>x)) is
#########  assuming data came from a lomax (when in fact it didn't)

## Mixture proportions ----------------------------------------
# We assume four categories of individuals with increasing movement.
pis <- c( 0.1, 0.2, 0.3, 0.4)

pars <- desired_mean_sd(mean = c(160, 300, 600, 1000), sd = c(90, 120, 160, 200))

mus <- pars$mu
sigsqs <- pars$sigsq
dens_cols <- c("#264653", "#2a9d8f", "#f4a261", "#e76f51")

### PLOT the mixture components -------------------------------
lnorm_densities <- purrr::map(1:4, function(y) stat_function(fun = dlnorm,
                                                             args = list(meanlog = mus[y], sdlog = sigsqs[y]),
                              color = dens_cols[y], size=1))
ggplot() +
  lnorm_densities +
  theme_minimal() +
  labs(y = "Density") +
  lims(x = c(0, 150)) -> densities_plot
densities_plot

```

```{r}

# Create the TRUE data

samp.size <- 50000
all.samples <- rep(0,samp.size)
categ <- rep(0,samp.size)

for(i in 1:samp.size){

  which.cat <- sample(1:4, size=1, replace=TRUE, prob=pis)
  all.samples[i] <- rlnorm(n=1,meanlog=mus[which.cat], sdlog=sigsqs[which.cat])
  categ[i] <- which.cat
}

sim.df <- data.frame(dists= all.samples, categ=categ)
```

```{r}
# Now the experiments and sampling

samp.sizes <- c(80, 200, 500, 800, 1000, 1600)
num.ns <- length(samp.sizes)
q.tests <- c(100, 250,500,1000)
num.qs <- length(q.tests)

all.sampsizes <- rep(samp.sizes,each=num.qs)
all.qtests <- rep(q.tests,num.ns)
ntests <- length(all.sampsizes)

cdfs.hat <- rep(0, ntests)

for(i in 1:ntests){

      ith.n   <- all.sampsizes[i]
      ith.samples <- data.frame(x = sample(all.samples, ith.n))
      mod1 <- lomax.glm(formula = ~1, my.dataf = ith.samples, response = ith.samples$x)
      ith.a <- mod1$alphas.hat[1]
      ith.k <- mod1$k.hat

      ith.q <- all.qtests[i]
      ith.cdf <- 1- lomax.cdf(x = ith.q, alpha = ith.a, k = ith.k)
      cdfs.hat[i] <- ith.cdf

}

true.cdfs <- rep(0,num.qs)
for(i in 1:num.qs){

  iq <- q.tests[i]

  true.cdfs[i] <- sum(all.samples>iq)/length(all.samples)

}

all.true.cdfs <- rep(true.cdfs,num.ns)

sim.test.df <- data.frame(all.sampsizes=all.sampsizes, all.qtests=all.qtests,cdfs.hat=cdfs.hat,
                          true.cdfs = all.true.cdfs)

```