### 2023-08-07 README.Rmd
Somehow, my previous README was deleted...

---
output: github_document
---

[![DOI](https://zenodo.org/badge/292314570.svg)](https://zenodo.org/badge/latestdoi/292314570)

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  eval = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# aracari <img src="images/aracari.png" align="right" height="139" />

<!-- badges: start -->
<!-- badges: end -->

My dissertation work is focused on the ecological consequences of individual variation in animal movement, from a modeling perspective. I am sharing this whole repository in an effort to keep science open and reproducible. All of my analyses and data are included in this repository. As of today (March 11, 2022) all three chapters are in the process of editing with coauthors for submission to scientific journals. Therefore, everything available on this repository is still a draft/manuscript/or work in progress. Please reach out if you have any questions or inquiries. I do ask that if you use any of the information within this repo, that you cite it. Thank you!

- CH1 - Consequences of incorporating heterogeneity in animal movement rates into models of seed dispersal. In this chapter we used a very simple random walk to model seed dispersal by animals in a homogeneous landscape. The main question here was whether or not incorporating individual variation in our simulations had an effect on the resulting dispersal kernels. We found that it does, and that by incorporating this variation, we allow for dispersal kernels to have fatter tails and increase the number of long-distance seed dispersal events. You can find the [Full Text](https://javirudolph.github.io/aracari/Ch1_movement_rates/FJRudolph_Ch1_0311.pdf) or the [Figures](https://javirudolph.github.io/aracari/Ch1_movement_rates/Ch1_Figures.html) and an [Appendix](https://javirudolph.github.io/aracari/Ch1_movement_rates/Ch1_Appendix_A.html)

- CH2 - Velocity-Based Analysis of Radio Telemetry Data to Explore Individual Variation in Animal Movement [Full Text](https://javirudolph.github.io/aracari/Ch2_distributions/FJRudolph_CH2_0311.pdf) [Figures](https://javirudolph.github.io/aracari/Ch2_distributions/Ch2_Figures.html)

- CH3 - Estimating Extreme Displacements Under a Lomax Generalized Linear Model Framework: The Role of Sample Size and Threshold Choice. [Full Text](https://javirudolph.github.io/aracari/Ch3_samplesize/FJ_Rudolph_CH3_0303.pdf).

-----